\documentclass[twoside]{../texlocal/linfoem}
\c\usepackage[dvipdfm]{hyperref}
\usepackage{epsf,../texlocal/refman2,../texlocal/tpage}
\usepackage{../texlocal/fileinclude}
\usepackage{../texlocal/urlsimple}
\c -*-latexinfo-*-
\c \anltmtrue
\anltmfalse
\c\pagestyle{empty}
\c
\c We really need a ``usage'' macro for the routines and macros
\def\usage#1{}
\c
\c the definition of \file should be changed so that it adds an
\c \penalty 1000\hskip 0pt +1fil \penalty -10000
\c before the actual text (allowing a line that the file name would stick
\c out of to instead be raggedright, a not perfect but better approach).
\c
\c See /home/beumer/GROPP/TMS/tohtml.tex
\c
\let\DOESupport\DefaultSupport

\def\tie{\nobreak\ \nobreak}
\def\bw{{\tt\char`\\}}
\textheight=9in
\textwidth=6.5in
\oddsidemargin=0in
\evensidemargin=0in
\topmargin=-0.5in
\hsize=\textwidth

\begin{document}
\bibliographystyle{plain}
\newindex{fn}

\begin{ifinfo}
\c %**start of header
\title{Users Manual for tohtml:\\
Producing True Hypertext Documents from LaTeX}
\author{William Gropp\\
Mathematics and Computer Science Division\\
Argonne National Laboratory}

\date{\today}

\c %**end of header

\maketitle
\end{ifinfo}

\begin{iftex}

\ANLTMTitle{Users Manual for tohtml:\\
Producing True Hypertext Documents from LaTeX}{\em William Gropp\\
Mathematics and Computer Science Division}{207}{March 1995}

\end{iftex}

\clearpage

\c \vskip 0pt plus 1filll
\c Copyright \copyright{} 1994 Argonne National Laboratory
\c
\c \clearpage
\c \pagenumbering{roman}
\c \setcounter{page}{3}
\c \pagestyle{plain}
\c \tableofcontents
\c \clearpage

\pagenumbering{arabic}
\pagestyle{plain}
\thispagestyle{plain}

\setfilename{../tohtml.info}

\node Top,        Introduction, (dir), (dir)
\c    nodename,   next,          previous, up

\c ---------------------------------------------------------------------------
\begin{iftex}

\pagenumbering{arabic}
\setcounter{page}{1}
\c \thispagestyle{empty}

\begin{center}
{\large\bf Users Manual for tohtml:\\
Producing True Hypertext Documents from LaTeX\\}
\vspace{.2in}
by \\
\vspace{.2in}
{\it
William Gropp\\
}
\end{center}

\vspace{.2in}

\addcontentsline{toc}{section}{Abstract}
\begin{abstract}
\noindent 
The World Wide Web has made it possible to use and disseminate
documents 
as ``hypertext.''  One of the major advantages of hypertext over conventional
text is that references to other documents or items can be linked directly
into the document, allowing the easy retrieval of related information.  A
collection of documents can also be read this way, jumping from one document
to another based on the interests of the reader.  This does require that the
hypertext documents be extensively cross-linked.
Unfortunately, most existing documents are designed as
linear documents.  Even worse, most authors still think of documents as linear
structures, to be read from front to back.  To deal with this situation, a
number of 
tools have been created that take documents in an existing word-processing
system or markup language and generate ``HTML,'' the hypertext markup language
used on the Web. 
While this process makes a single document available in a convenient form on
the Web, 
it does not give access to cross-document linking, a major advantage of
hypertext.  This manual describes a program, {\tt tohtml}, that takes LaTeX
input 
files, as well as files of link information, and produces a hypertext
document that can contain extensive cross-links.  A related program, {\tt
doctext}, aids in the generation of manual pages that can be referenced by a
LaTeX document.

\end{abstract}
\end{iftex}


\clearpage
\c ---------------------------------------------------------------------------



\c ---------------------------------------------------------------------------

\node Introduction,tohtml,Top,Top
\section{Introduction}
\label{chap-intro}
Generating true hypertext documentation is a task that requires skill and hard
work.  The tools described in this report are intended to help you create
simple hypertext documents that include multimedia elements as well as
extensive cross-referencing, without requiring you to learn a new
word-processing system (as long as you already know LaTeX \cite{Lamport86}).
The documents 
produced by this system, while relatively simple, do include hot links to
other documents, the ability to link individual names (such as routine or
person names) to other documents, and simple multimedia elements.

Specifically, we describe the program \code{tohtml}, which converts LaTeX
files to HTML (the Web hypertext markup language).  This C program is fast and
can handle very large documents.  The program \code{doctext} can be used to
generate manual pages for routines written in C that can be referenced
automatically by documents built with \code{tohtml}.  Finally, a LaTeX style
file, \file{tools.core/info2rtf/anlhtext.sty} provides a way to use LaTeX
and include hypertext links, graphics, and movies into a document.

Other tools exist for translating LaTeX to HTML; perhaps the best known is
LaTeX2html.  This program has different strengths and weaknesses when compared
with \code{tohtml}; if you have trouble with one program, you should try the
other.

LaTeX in this document refers to the original LaTeX, not the new and
incompatible LaTeX2e.  Support for LaTeX2e may be provided in the future.

\node tohtml,,,Top
\section{tohtml}
The \code{tohtml} program is a C program that can read LaTeX files and
generate HTML files.  Command-line options give you control over the output,
including the way that the document is split into separate HTML files and the way
complicated expressions are handled.  

\subsection{Getting Started}
Using \code{tohtml} is easy.  For example, if you wish to convert a LaTeX file
named \file{mypaper.tex} into an HTML document, you can do
\begin{example}
tohtml -default mypaper.tex
tohtml -default mypaper.tex
\end{example}
Note that, just as you normally need to run LaTeX twice to get all of the
references correct, you also need to run \code{tohtml} twice.  The
command-line option \code{-default} provides good basic choice of the many
options provided by \code{tohtml}.  The first run will also generate a number
of error messages (such as \code{Could not find Introduction in topicctx} and 
\verb+\ref{fig:upshot1} unknown (mypaper.tex line 792)+); the second run
should generate no error messages if all goes well.

The output will be left in a subdirectory with name \file{mypaper}; you can
view the result with \code{xmosaic mypaper/mypaper.html}.

One important note:  for \code{tohtml} to handle bibliography entries
correctly, you should keep the files that LaTeX and BibTeX generate 
(e.g., the \file{mypaper.bbl} file).  

\subsection{Command Line Arguments}
To use \code{tohtml}, you give it the name of the LaTeX
file to process:
\begin{example}
tohtml foo.tex
\end{example}
Command-line options to \code{tohtml} allow you to change the behavior.  The
option \code{-default} sets a common selection of options:
\begin{example}
tohtml -default foo.tex
\end{example}

A complete list of the command-line options follows.  Some of these (e.g.,
\code{-mapref}) are discussed in more detail later.

The following options control how the document is divided into individual Web
pages.
\begin{description}
\item[-nocontents]Suppress the generation of the contents page (which has
links to all of the other sections).  This is useful for short documents.

\item[-split n]Split the document into separate files, based on the sectioning
commands (e.g., \verb+\chapter+, \verb+\section+).
The value of \code{n} indicates where to begin breaking the file: a value of
\code{0} puts chapters into separate files, a value of \code{2} puts
subsections into a separate file, and a value of \code{-1} forces the entire
document into a single file (note that any image or graphics files are still
separate).
\end{description}

The following options control how \code{tohtml} handles various kinds of LaTeX
commands.
\begin{description}
\item[-citeprefix str]
\item[-citesuffix str]
Set the prefix and suffix strings used around
\verb+\cite+ in text.  The default values are \code{[} and \code{]}.

\item[-cvtlatex]Convert LaTeX that cannot be represented directly by HTML
into bitmaps that 
are included in the document.  If this option is not used, the text is
processed 
as is, with all LaTeX commands removed and text (and arguments to the
commands) unchanged.

\item[-cvttables]Convert tabular environments into bitmaps.  This produces
nicer tables than the default (which attempts to line up columns) but can
produce large bitmaps that some systems may not be able to handle.

\item[-cvtmath]Convert math and display math environments into bitmaps.  This
produces nicer representation of the formatted mathematics, but can produce
large bitmaps.

\item[-iftex]Include text in \verb+\begin{iftex}+ to
\verb+\end{iftex}+ (for LaTeXInfo documents).

\item[-simplemath]Use italics for
expressions in LaTeX math mode that do not involve any special characters.
For example, \code{$3$} would be changed into an italic 3, rather than a small
bitmap.  Not yet available.

\item[-default]Set a common set of options.  Equivalent to \code{-cvtlatex
-cvttables -cvtmath -iftex -split 2 -useimg}.

\item[-basedef filename]Read filename for additional definitions.  This lets
you define some TeX and LaTeX commands as having a specific behavior when
generating HTML files (\pxref{User-defined Replacements}).

\item[-userpath pathlist]For each \code{documentstyle} option or
  \code{usepackage} name, look in \code{pathlist} for a file of the form
  \code{name.def}.  For example, \verb+\usepackage{refman}+ will cause
  \code{tohtml} to search for \code{refman.def} in \code{pathlist}.  If
  \code{-userpath} is not given, no files will be searched for.  The
  environment variable \code{DOCTEXT_USERPATH} may be used instead of
  \code{-userpath}. 
\end{description}


The following options control some aspects of the layout of each page,
particularly the 
presence of the navigation buttons to other pages.
\begin{description}
\item[-notopnames]Suppress generation of the links at the top of the
page.  This is useful when generating a single-page document, in combination
with \code{-split -1}.

\item[-nonavnames]Suppress generation of the links at the bottom of the
page.  This is useful when the sections in the document are short.

\item[-nobottomnav]Suppress generation of the links and buttoms at the
bottom of each section.

\item[-beginpage filename]Add the HTML in \code{filename} to the top of every
generated HTML page.

\item[-endpage filename]Add the HTML in \code{filename} to the bottom of every
generated HTML page.  This is particularly useful for adding links to indexes
and home pages to a document.
\end{description}


The following command-line arguments control the generation of links to other
documents. 
\begin{description}

\item[-mapref filename]Filename contains a list of mappings from citation keys
(the name in a \verb+\cite+ command) to HTML links (\pxref{Building a Map
File}). 

\item[-mapman filename]Filename contains a list of mappings from tokens
(currently defined as sequences of letters only) to HTML links  
(\pxref{Building a Map File}).
\end{description}

These commands control details of the generated HTML file.

\begin{description}
\item[-gaudy]Use images of colored balls for
bullets in itemized lists.

\item[-useimg]Instead of generating the bitmap, uses a file that has the name
that would be used for that bitmap.  This
is useful when making the second run of \code{tohtml}.  It is meaningful only when
\code{-cvtlatex} is used.

\item[-basedir dirname]Add an HTML base command to the main file.
For example, \code{-basedir "http://www.mcs.anl.gov/mpi"} causes
\begin{example}
<base href="http://www.mcs.anl.gov/mpi">
\end{example}
to be written to the file.

\end{description}

The following commands are used for debugging the behavior of tohtml itself.
\begin{description}
\item[-debugout]Write out information about the generation of output
\item[-debugdef]
\item[-debugscan]
\item[-debugfile]
\end{description}

\node doctext,,,Top
\section{doctext}
One of the most difficult tasks in creating extensive hypertext is generating
the initial documents and providing an easy way to link to them.  The
\code{doctext} \cite{doctext} program can be used to generate versions of
Unix-style man 
pages from the C source code of routines.  This program can generate
\code{nroff} (for using \code{man} and \code{xman}), LaTeX (for generating
printed manuals), and HTML.  For the HTML to be useful, there must be an easy
way to create links to the generate documents.  This section describes how to
do that; \code{doctext} itself is documented in \cite{doctext}.

To generate HTML man pages of a collection of source files in
\file{/home/me/foo}, 
do the following:
\begin{verbatim}
cd
mkdir www
mkdir www/man3
cd foo
doctext -html -index ../foo.cit \
        -indexdir http://www.mcs.anl.gov/me/foo/www/man3 \
        -mpath ../www/man3 *.[ch]
cd ..
\end{verbatim}
This puts the HTML files into the directory \file{www/man3} and the index (in
the correct format for \code{-mapman}) into file \file{foo.cit}.  Once you are
sure that the files are correct, you can move them into the Web area with
\begin{example}
cp -r www /mcs/www/home/me
\end{example}
(assuming that \file{/mcs/www} corresponds to \code{http://www.mcs.anl.gov} in
the \code{-indexdir} argument).

To generate an HTML listing of the routines, you can execute the following
script, with, of course, the appropriate text:
\begin{verbatim}
#! /bin/sh
echo '<TITLE>Web pages for My Routines</TITLE>' >> www/index.html
echo '<H1>Web pages for My Routines</H1>' >> www/index.html
echo '<H2>My Routines</H2>'  >> www/index.html
echo '<MENU>' >> www/index.html
ls -1 www/man3 | \
    sed -e 's%^\(.*\).html$$%<LI><A HREF="man3/\1.html">\1</A>%g' \
        >> www/index.html
echo '</MENU>' >> www/index.html
\end{verbatim}
If you have only a few routines to document, you can dispense with the second
directory level above (the \file{man3}).  However, you might find it valuable
to follow (at least loosely) the Unix man-page format, with commands and
installation instructions in \file{man1} and routines spread across
\file{man1} to \file{man8}.

\node Building a Map File, , , Top
\section{Building a Map File}
One of the most useful features of \code{tohtml} is its ability to
automatically replace citations and names with hot links into other documents.
This is done by providing one or more \code{-mapref} files (for citations)
and \code{-mapman} files (for names, currently defined as strings consisting
only of letters).
Currently, the format of this file is fairly ugly but easy to use:
\begin{quote}
\code{cite:+}{\em cite-name}\code{++}{\em cite-text}\code{++++}{\em kind}\code{+}{\em URL}
\end{quote}
where the items are as follows.
\begin{description}
\item[cite-name]A citation key that is used in a LaTeX \verb+\cite+
command, such as \verb+\cite{tohtml-user-ref}+.

\item[cite-text]The name that should replace the citation use.  For
example, if the document contains a \verb+\cite{tohtml-user-ref}+, and you
wish this to turn into a hot link with text \code{[Tohtml User Manual]},
you would use \code{Tohtml User Manual} for {\em cite-text}.

\item[kind]The field is used to indicate the type of the reference.  This is
not yet implemented.  Use the value \code{manual} for user manuals and
\code{cite} for everything else for the time being.

\item[URL]The URL (Universal or Uniform Resource Locator) that the item will
link to.  These are paths such as
\code{http://www.mcs.anl.gov/home/gropp/tohtml/tohtml.html}.

\end{description}

All of these items must be on a single line; the exact number of \code{+}
characters must be used.

You can comment out a line in a mapref or mapman file by placing a ``\code{;}'' in
the {\em first} column. 

Here is a sample map file for citations.  Note that \code{\bw{}cite\{gropp\}}
will 
turn into \code{gropp@mcs.anl.gov} in the text, with a link to the Web home
page for gropp.

\begin{tiny}
\begin{example}
cite:+sles-user-ref++SLES User Manual++++manual+http://www.mcs.anl.gov/petsc/solvers.html
cite:+gropp++gropp@mcs.anl.gov++++person+http://www.mcs.anl.gov/home/gropp/index.html
cite:+petsc-man-pages++PETSc Manual Pages++++manual+http://www.mcs.anl.gov/petsc/ref/index.html

\end{example}
\end{tiny}

\subsection{Map Files for man Pages}
Building a map file for a list of man pages generated by, for example,
\code{doctext}, is a daunting prospect.  Fortunately, the map file can be
generated automatically when the man pages are created if \code{doctext} is
used. 
To do this, add the options
\begin{quote}
\code{-index} {\em indexfile} \code{-indexdir} {\em indexdir}
\end{quote}
to the \code{doctext} command.  The {\em indexfile} is a file to which entries
in the proper form for \code{-mapman} will be {\em appended}.
The value of {\em indexdir} is the root directory for the generated man pages.
For example, to use \file{http:/www.mcs.anl.gov/home/gropp/man} as the root
and the file \file{manref.cit} to hold the \code{-mapman} references, use
\begin{example}
doctext -index foo.cit \bw
    -indexdir "http:/www.mcs.anl.gov/home/gropp/man" ...
\end{example}

\node Fine-Tuning,,,Top
\section{Fine-Tuning}

This section discusses how to customize the appearance of the pages generated
by \code{tohtml}, how to interpret error messages, and how to extend the
translations that \code{tohtml} recognizes.

\node The anlhtext style file, , , Fine tuning
\subsection{The anlhtext Style File}
One way to fine-tune a document is to use a special LaTeX style file that
contains LaTeX commands that have special meaning to \code{tohtml}.
This makes it possible to maintain a single document for both hardcopy and
hypertext versions of a document.

\begin{description}
\item[URL]Gives the URL for a document.  In LaTeX, it is set in the current
font; in HTML, it becomes an active link whose text is the link name.
Usage is \verb+\URL{url-text}+.
\item[AURL]Gives the URL with a different link text.  In LaTeX, gives the URL
in the \verb+\tt+ font, followed by the link text.  In HTML, just the link
text is displayed; clicking the link text jumps to the URL.  Usage is 
\verb+\AURL{url}{link text}+.
\item[hcite]Includes a link to a URL that appears {\em only} in the
HTML document.  Usage is \verb+\hcite{url}+.
\item[hcitea]Includes a link to a URL with a specified text string; the
URL is 
used only in the HTML version.  Usage is \verb+\hcitea{url}{text string}+.
\item[href]Refers to a section or subsection by title.  For example,
if you used \verb+\subsection{A sample}+, then \verb+\href{A sample}+ would
provide a link to that subsection in the HTML.  In LaTeX, it sets its argument
in the current font.  Usage is \verb+\href{section title}+.
\end{description}

The following have not yet been implemented.  Comments are welcome.
\begin{description}
\item[htmlhr]Generates an {\em hrule} in the HTML.  Ignored in LaTeX.
\item[html envirionment]\verb+\begin{html}+ to \verb+\end{html}+ may be used
to imbed HTML commands directly; in LaTeX, no output will be generated.
\item[movie]Includes a reference to an MPEG movie file with a given icon.
\item[sound]Includes a reference to a sound file with the generic sound icon.
\item[graphic]Includes an image (gif) file.
\end{description}

\node  Adding Pointers to Other Pages to Every Page, , , Fine-tuning
\subsection{Adding Pointers to Other Pages to Every Page}

In many situations, you would like to add pointers to the bottom of every
page, for example, a pointer back to a home page.  With \code{tohtml} this is
very easy.  First, create a file, say \file{bottom.html}, that contains the
appropriate HTML code:
\begin{small}
\begin{example}
Return to <A HREF="node179.html">MPI Standard Index</A><BR>
Return to <A HREF="http://www.mcs.anl.gov/mpi/index.html">MPI home page</A><BR>
\end{example}
\end{small}
The first line provides a return to a document in the current directory, in
this case, the index of the MPI Standard.  
The second line provides a return to the MPI home page; this will work from
any location.

To have these incorporated into the output of \code{tohtml}, use the option
\code{-endpage}:
\begin{example}
tohtml -endpage bottom.html ...
\end{example}

\node Adding a Disclaimer to the Top of Every Page, , , Fine tuning
\subsection{Adding a disclaimer to the top of every page}
Sometimes, you may wish to add text to the top of every generated page.
You can do this by putting the text into a file, say \file{top.html}, and then
using the \code{-beginpage} option of \code{tohtml}:
\begin{example}
tohtml -beginpage top.html ...
\end{example}
Here \file{top.html} might contain
\begin{example}
This material is a draft and should not be quoted<BR><HR>
\end{example}

\node Errors, , , Fine tuning
\subsection{Errors}
Errors are reported in two places.  Some messages go to standard error. Lists
of LaTeX commands unknown to \code{tohtml} or citations without hyperlinks are
written into the file \file{latex.err}.  In addition, \code{tohtml} will not
let you redefine some basic LaTeX commands such as \verb+\section+; normally,
this is the behavior that you'd want anyway.

\node User-defined Replacements, , , Fine-tuning
\subsection{User-defined Replacements}
You can provide your own definition of many TeX and LaTeX commands by
including a {\em definitions} file.  This file contains entries of the form
\begin{quote}
{\em operation} {\em name} {\em optional-string}
\end{quote}
Lines may be commented out by starting them with the character \code{#}.
\begin{description}
\item[operation]This is one of \code{nop} (no operation and remove arguments),
\code{dimen} (TeX dimension) \code{asis} (no operation but leave arguments),
\code{name} (replace with a string), or \code{raw} (replace with verbatim
text, such as an HTML command), and \code{exec} (insert TeX commands).

\item[name]This is the name of the TeX or LaTeX command, without the initial
backslash.  That is, to redefine \verb+\samepage+, use the name
\code{samepage}.

\item[optional-string]This is null (for \code{dimen}), an integer for
the number of arguments (for \code{nop} or \code{asis}), or the replacement
string (for \code{name}).  Currently, there is no way to specify leading or
trailing spaces.

\end{description}

For example, when \code{tohtml} starts, it reads a file that contains the
definitions of some basic commands, such as
\begin{example}
dimen topsep
asis underline 1
nop thispagestyle 1
name mu 0 u
raw seprule 0 <HR>
\end{example}
The first line states that \verb+\topsep+ is a dimension; \code{tohtml} will
understand what to do when it sees this name.  The second line says that
\verb+\underline+ is a command with a single argument and that the
appropriate HTML for this is to use the argument (thus, \verb+\underline{abc}+
will become \code{abc} in the HTML document).
The third line says that \verb+\thispagestyle+ is a command with a 
single argument and can be ignored (performs no operation, or `no-op').
The fourth line tells \code{tohtml} to replace \verb+\mu+ with \code{u}.  
The last line tells \code{tohtml} to replace \verb+\seprule+ with the HTML
command \verb+<HR>+.
\c The
\c last line shows the use of quotes to imbed a space in a name; this causes 
\c the
\c command `\verb+\ +' to generate a single space in the HTML output.

Any number of \code{-basedef} files may be specified; they are processed in
order starting from the left.  The default basedef file is always read first.

\node Examples, , , Top
\section{Examples}
This section gives examples of using \code{tohtml} for various kinds of
documents.

\subsection{Single-Page Descriptions}
Sometimes, you wish to use LaTeX sectioning commands to generate headings in
the HTML file without generating separate HTML files, for example, when each
section is very short.  This example shows how to accomplish this.
In addition, it
\begin{itemize}
\item adds some HTML to the bottom of the page (\code{-endpage
petsc_end.html}),
\item suppresses the contents list (\code{-nocontents}),
\item suppresses the links to jump between sections (\code{-notopnames
-nonavnames}),
\item makes the use of \verb+\cite+ {\em not} place brackets around the
reference (\code{-citeprefix "" -citesuffix "" }),
\item automatically converts citations to links to other documents
(\code{-mapref ../../html/petsc.cit}), and
\item places the output into a separate directory (\code{split 0} instead of
\code{split -1}).
\end{itemize}

\begin{verbatim}
tohtml -mapref ../../html/petsc.cit -endpage petsc_end.html \
        -nocontents -notopnames -nonavnames -split 0 \
        -citeprefix "" -citesuffix "" domain.tex
tohtml -mapref ../../html/petsc.cit -endpage petsc_end.html \
        -nocontents -notopnames -nonavnames -split 0 \
        -citeprefix "" -citesuffix "" domain.tex
\end{verbatim}
In addition, the option \code{-gaudy} may be used to add color to any lists.

\subsection{Chosing LaTeX executable}
Because of the introduction of LaTeX2e, which is often installed as just
\code{latex}, it may be necessary to select a particular \code{latex}
executable.  You can do this with the environment variable
\code{TOHTML_LATEX}; just set that to the path needed for LaTeX.
For example,
\begin{verbatim}
    setenv TOHTML_LATEX /usr/local/bin/latex
\end{verbatim}

\subsection{Books}
A major advantage of \code{tohtml} is its ability to handle very large
documents.  Here are the commands that are used to process the MPI Standard:
\begin{verbatim}
tohtml -default -endpage ../mpi-tail.html mpi-report.tex
tohtml -default -endpage ../mpi-tail.html mpi-report.tex
\end{verbatim}
The \code{-endpage ../mpi-tail.html} option is used to add pointers to the
index of the MPI standard and the MPI home page to every page.

\section{Comments, Bugs, and Suggestions}
Please send any comments to \code{gropp@mcs.anl.gov}.

\begin{tex}
\addcontentsline{toc}{section}{References}
\bibliography{../tools}
\end{tex}

\end{document}
